Busch, Khakhalin. Intrinsic temporal tuning

### 2019-04-19 reviews and responses.md

> - There is also an inconsistent use terminology. For example, “spikiness” and
“excitability” are sometimes used in attempt to describe distinct cell characteristics,
but sometimes it seems like they are being used interchangeably. This makes
interpreting the written results difficult to follow.

In this updated version, we consistently use "spikiness" as a name for the variable, for the first time introduced in Figure 2C. For earlier results section, we replaced "spikiness" with "average number of spikes", and in all instances when we talked about the general idea of neuronal excitability, we called it neuronal excitability. As the only exception to this rule, we opted for the word "spikness" in the first paragraph of Introduction, to prime the reader for the use of this word later in the text.

> - “Compared to control, neurons from animals exposed to visual stimuli had a flatter
amplitude tuning curve (Fig 1D,E).” What is the reader supposed to be looking at in
the figure? Is this referring to larger conductances for a fixed duration? I think this is
what the authors’ mean, but more intuitive descriptions that matches the figures or
vice versa is needed to make the whole results section easier to digest.

We improved the sentence:

Compared to control, neurons from animals exposed to visual stimuli had a flatter amplitude tuning curve (**less of a difference between columns of Figure 1D, or different curves in each panel of Figure 1E**), as they did not increase their spiking as fast in response to larger conductances

@@

We indeed received some conflicting advice in the past about the level of detail necessary to communicate this message: for example, people who have studied Simpson's paradox in their statistics classes had found our use of this term very helpful, while people who haven't heard it before, found it distracting. For now, we opted to include it, but we are aware that what makes the text more relatable to one group of researchers may make it less transparent for other groups.

Nevertheless, in current version of the paper, we shortened this section from 411 to 294 words.

> - In reference to the text at the bottom of page10: Is it possible to include a summary
figure of data from the stepwise generalized linear model in Fig. 4? The purpose of
this part of the study was to try and identify electrophysiological mechanism(s) that
explain the tuning variables. The heart of Fig. 4 should be that they don’t, and figure
that captures this would be helpful.

TODO: How to make a figure like that? They refer to a text about the stepwise generalized linear model. Is there a good visualization here?

### 2019-05-31 main.tex
Finished all responses + Discussion improved

It is often assumed, in fields as diverse as connectomics and machine learning, that the main difference between functional and dysfunctional neural networks lies in their connectivity \citep{takemura2014,hildebrand2017,bassett2017network,reimann2017}. However, biological neurons also demonstrate complex and multifaceted intrinsic tuning  \citep{oleary2013,evans2015}, with cells within a single network having different activation thresholds \citep{kole2012}, bursting behaviors \citep{popovic2011}, and inactivation profiles \citep{bianchi2012}. Dysregulation of intrinsic plasticity affects the dynamics of networks  \citep{tien2018}, and can lead to a loss of function \citep{marcelin2009}. And yet, with the possible exception of oscillatory networks \citep{marder2011,picton2018control}, the exact ways in which intrinsic plasticity contributes to network tuning, remain unclear. We know that neurons adjust their spikiness to match the levels of synaptic activation they experience \citep{aizenman2003,titley2017}, but we also know that biophysically, they often have the means to become tuned to different temporal patterns of activation \citep{branco2010dendritic,jarvis2018morphology,ohtsuki2018bursting}. Do they rely on this type of tuning in practice?

In this paper, we investigate whether exposure to sensory stimuli with different temporal dynamics would change intrinsic temporal tuning of sensory neurons in the optic tectum of \textit{Xenopus} tadpoles. Tadpole tectum is an ideal model for exploring this question: it is a highly malleable distributed network of about 10\textsuperscript{4} neurons \citep{pratt2013}, involved in stimulus discrimination and sensorimotor transformations \citep{dong2009,khakhalin2014}. In development, tectal neurons acquire diverse phenotypes, that are then shaped by sensory experiences \citep{xu2011,ciarleglio2015}. Circuits in the tectum can learn and reproduce temporal patterns to which they were exposed \citep{pratt2008}: a property that could in principle be achieved through synaptic changes alone \citep{lukovsevivcius2009}, but which may be easier to achieve through intrinsic temporal tuning \citep{narayanan2008,beatty2014}. Finally, tectal neurons exhibit strong Na channel inactivation, which seems to play a role in collision detection \citep{jang2016}, and is a known target for intrinsic plasticity mechanisms \citep{bianchi2012}.

Specifically, we ask two questions about the properties of intrinsic plasticity in tectal networks. First, we check whether intrinsic plasticity in response to sensory stimulation is limited to changes in overall excitability, or whether it also retunes neuronal sensitivity to faster and slower inputs. Second, we ask whether intrinsic temporal tuning of tectal cells is coordinated with the typical length of synaptic currents they receive, or whether two types of tuning are independent. In our previously published large-scale census of tectal cells \citep{ciarleglio2015}, we observed no interaction between intrinsic and synaptic phenotypes, and observed no signs of temporal tuning (ibid,  figures 2, 4). We argue however, that the current clamp protocol used in earlier studies was not adequate for the task, so in this paper we relied on a dynamic clamp technique.

The main benefit of the dynamic clamp approach, compared to commonly used current step injections, is that it allows a more realistic simulation neuronal response to synaptic conductances \citep{prinz2004}. In dynamic clamp, the electric current injected into the cell is dynamically adjusted, based on a predefined formula that depends on cell membrane potential and time. This means that with dynamic clamp recordings we can excite a neuron in a controlled manner, but still allow both its voltage and transmembrane current to change, preserving feedback interactions between active currents and transient voltage-gated channels inactivation \citep{ma1996dynamic}, which may be an important mechanism for intrinsic temporal tuning \citep{platkiewicz2011}. Also, as a second methodological innovation, instead of relying on one type of sensory stimulation, here we used five different stimulation protocols, and compared their effects on intrinsic tuning.

@@

First we asked whether our stimulation protocols caused any changes in intrinsic excitability of tectal neurons. From previous studies, we knew that in tadpoles exposed to four hours of LED flashes, tectal neurons became more excitable \citep{aizenman2003,ciarleglio2015}. However, the stimuli we used in the present study were weaker, and similar to those used in behavioral experiments \citep{khakhalin2014,james2015,truszkowski2017}. We presented a checkerboard pattern that inverted once a second either instantaneously (dubbed “Flash”; Figure 1C left), or with a slow transition over the course of a second (old black squares shrank to white, while new black squares grew from old white squares, dubbed “Looming”; Figure 1C right).

@@

After conditioning, we excised the brain, performed whole cell patch clamp on tectal neurons (Figure 1A), and counted spikes produced in dynamic clamp mode in response to simulated synaptic conductances of different durations and amplitudes (Figure 1B). We used conductances of 4 different durations (100, 200, 500, and 1000 ms), and 3 different amplitudes (peak conductances of 0.2, 0.5, and 1.0 nS), matching the range of synaptic currents observed in tectal circuits \textit{in vivo} (\citealt{xu2011,khakhalin2014,ciarleglio2015}; see Methods). Based on prior research, we expected neurons exposed to flashes (checkerboard inversions) to be more excitable, but contrary to our expectations, they were on average less spiky, and generated 0.4$\pm$0.4 spikes, across all types of dynamic clamp inputs, compared to 0.9$\pm$1.0 spikes in control (Figure 1D, left column; F(1,677)=30.4, p=5e-8, n cells=28, 29). (Here and below, F-values are reported for a multivariate fixed effects analysis of variance with selected interactions, where cell id is used as a repeated measures factor; see Methods for a detailed description). Neurons exposed to looming transitions also spiked less than control neurons (0.6$\pm$0.4 spikes; F(1,629)=9.6; p=0.002; n=28, 25), but more than those exposed to instantaneous "flashes" (F(1,641)=3.7; p=4e-6; n=29, 25). This was likewise unexpected, as in the tadpole tectum, looming stimuli are known to elicit stronger responses compared to instantaneous inversions \citep{khakhalin2014}, yet the effect yielded by looming stimuli was weaker. To sum up, both the expected sign of the effect, and the relative

We then mapped the amplitude tuning of neurons (amplitude transfer function, or gain), by looking at how an increase in transmembrane conductance translated into increased spike output. Compared to control, neurons from animals exposed to visual stimuli had a flatter (constant or hill-shaped, rather than increasing) amplitude tuning curve (Fig 1D,E), as they did not increase their spiking as fast in response to larger conductances (F(1,677)=15.9, p=8e-5; and F(1,629)=8.3, p=0.004 for flash- and looming transitions respectively). The flattening of response curves seemed slightly more pronounced for flashes than for looming, but this difference was not significant (F(1,641)=3.5; p=0.06).

These results show that prolonged stimulation had an effect on intrinsic excitability, but its direction was opposite to what was previously described \citep{aizenman2003,ciarleglio2015}, as neurons became less excitable. Moreover, while looming stimuli are known to be more salient, both behaviorally and physiologically \citep{khakhalin2014}, they had a weaker long-term effect on neuronal excitability in comparison to less salient flashes. We conjecture that the difference in the direction of change is due to our stimuli being weaker than those used in earlier studies (see Discussion), and we further explore the difference between flashes and looming stimuli below.

@@

We then examined whether different types of sensory activation would differentially reshape temporal intrinsic tuning in tectal neurons. As changes in intrinsic properties seemed homeostatic (increased activation led to reduced spiking), by the same logic, one could expect shorter stimuli (flashes) to selectively suppress responses to shorter synaptic inputs. Alternatively, shorter stimuli could reshape the network, making neurons better adjusted to working with short bursts of activation, as  previously described for synaptic processing \citep{aizenman2007} and recurrent activity in the tectum \citep{pratt2007,shen2011}.

Across input conductances of different lengths (100, 200, 500, and 1000 ms), neurons exposed to flashes and looming stimuli responded differently than in control (F(1, 677)=25.1; p=7e-7; and F(1,629)=12.0; p=6e-4 respectively). In control neurons, longer inputs typically evoked stronger spiking, whereas neurons from stimulated animals had flatter tuning curves, with a plateau, or even a decrease in spike number for longer conductance injections (Figure 1E). In essence, while control neurons preferred longer inputs, stimulated neurons developed a preference for shorter synaptic inputs, and this change was more pronounced in neurons exposed to flashes than in those exposed to looming stimuli (F(1,641)=7.5; p=0.006). This suggests that the change in overall intrinsic excitability, and the change in temporal tuning, follow two different kinds of logic. The overall excitability is homeostatic, as neurons became less excitable in response to stronger stimulation. The temporal retuning however can be better described as “adaptive”, as neurons exposed to shorter stimuli (flashes) became relatively \textit{more} responsive to shorter stimuli, and less responsive to longer stimuli, which is the opposite of what one would expect for a purely homeostatic retuning. We chose to call this type of plasticity "adaptive", as presumably it means that after exposure to faster stimuli, neurons become more equipped to process faster patterns of activation \citep{stemmler1999information}.

@@

While the optic tectum (homologous to superior colliculus in mammals) is often described as a primarily visual area, it is also involved in heavy multisensory computations \citep{stein2014}. In tadpoles, it integrates visual information with inputs from mechanosensory, auditory, and lateral line modalities \citep{deeg2009,pratt2009trigeminal,hiramoto2009,felch2016,truszkowski2017}, but the logic of this integration is still unclear. We wondered whether acoustic stimuli would reshape intrinsic properties of tectal neurons, and whether this retuning would be similar to that produced by visual stimuli.

To test this question, we exposed tadpoles to four hours of behaviorally salient “click” sounds \citep{james2015,truszkowski2017}, provided at the same frequency (every second) as visual stimuli in the first set of experiments. We found (Fig 1D, E) that exposure to startle-inducing sounds (group “Sound”) did not lead to significant changes in either average number of spikes (0.8$\pm$0.7 spikes; F(1,689)=1.7; p=0.2; n=28, 30), amplitude transfer function (F(1,689)=2.0; p=0.2), or temporal tuning curve (F(1,689)=2.0; p=0.2). This suggests that acoustic stimuli did not strongly activate tectal circuits during conditioning, despite being more behaviorally salient (for our stimuli, at the onset of stimulation, acoustic clicks evoked startle responses in about 50-80\% of cases, compared to 5-10\% for checkerboard inversions \citep{james2015,truszkowski2017}). This was not necessarily surprising, as auditory and mechanosensory inputs are known to have different cellular \citep{pratt2009trigeminal,felch2016} and subcellular \citep{hiramoto2009} targets in the tectum, and thus can recruit tectal networks to a different extent.

We then combined visual and acoustic stimuli in two different ways and looked at the effects of multisensory stimulation on the intrinsic properties of tectal neurons. For some animals, we synchronized the instantaneous checkerboard inversions (flashes) with sound clicks (dubbed “Sync”), while for others we staggered visual and acoustic stimuli by half a period (500 ms; dubbed “Async”). We found (Fig 1D, E) that, after four hours of multisensory stimulation, tectal neurons were more excitable than after visual stimulation alone (0.6$\pm$0.4 spikes, F(1,689)=11.2, p=8e-4; and 0.7$\pm$0.6 spikes, F(1,665)=41.7, p=2e-10, for sync and async respectively, across all testing conditions). However, compared to the “Flash” group, the tuning curves in multisensory groups were less flat, with a stronger effect in the Async group (for amplitude tuning: F(1,689)=1.3, p=0.3, and F(1,665)=2.4, p=0.02 in Sync and Async groups respectively; for temporal tuning: F(1,689)=9.3, p=0.002, and F(1,665)=22.8, p=2e-6 respectively). This suggests that on their own sound clicks had little effect on tectal excitability, but when added to visual flashes, sound clicks negated effects of retuning that visual stimulation would have had (Fig 2C).

% Make this paragraph above more clear, lead the reader throught the tension. Also, add an interpretation below, and there say that it's either inhibition, or changes at the level of individual dendritic trees. Make it clear that this result WAS surprising. And if it's dendritic processing, then explain how it may help.

@@

% Citations that may be used, when talking about inhibition: and may also differentially recruit tectal inhibitory circuits \citep{liu2016,hamodi2016}.

% Also reference Deeg with her independent intrinsic tuning (in the same neuron!) to visual and mechanosensory inputs (does it suggest a dendritic tree?)

\begin{figure*}[!t]
\includegraphics[width=\linewidth]{fig2.pdf}
\caption{
Quantification of changes in temporal tuning in response to sensory experience. (\textbf{A}). An illustration of how “Temporal tuning” and “Amplitude tuning” values were calculated. For the temporal tuning measure, the value of zero corresponds to linear dependency (blue line), positive values - to an accelerating, supralinear curve (red), and negative values - to a plateau-shaped curve (purple, yellow). For amplitude tuning, higher values correspond to faster increase in spiking with increased conductance. (\textbf{B}). Amplitude tuning of neurons across different experimental groups (stars show t-test p$<$0.05 compared to control). (\textbf{C}). Temporal tuning and average spikiness of neurons in different experimental groups; in each plot all neurons across all groups are shown in gray, while neurons from one target group are shown in color; means are shown as black dots; ellipses represent 95\% normal confidence regions. Two outliers (top right corner) are brought within the axes limits. (\textbf{D}). Same data as in (C), shown as averages for each group, with 95\% confidence intervals. Black arrows show the effects of sound clicks, when they were added to control, and when they were added to “Flashes” to form two types of multisensory stimuli.}
\end{figure*}

To visualize and interpret differences in neuronal tuning, we quantified each of the three aspects of intrinsic tuning (average spikiness, amplitude tuning, and temporal tuning) with one value per neuron (see Methods). We used the mean number of spikes across all conditions as the measure of "spikiness"; the linear slope of the number of spikes as a function of conductance amplitude as the measure of "amplitude tuning", and the quadratic term of the curvilinear regression for the number of spikes as a function of input duration as the value characterising the "temporal tuning" or adaptation index of each neuron (Figure 2A). The numerical values of these "tuning coefficients" are not easily interpretable, but they capture the character of tuning curves for each neuron (Figure 2A). All three parameters differed across experimental groups: F(5,160)=3.1, p=0.01 for average spikiness (Figure 2C; see Methods for model description); F(5,160)=4.8, p=4e-4 for amplitude tuning (Figure 2B); and F(5,160)=3.6, p=4e-3 for temporal tuning (Figure 2C).

A visual comparison of neuronal tuning in different experimental groups (Fig 2C) shows that acoustic stimulation had opposite effects when provided on its own (without visual stimulation) than when added to visual flashes. Compared to control neurons, cells exposed to sound had slightly lower amplitude tuning coefficients (flatter curve, Cohen’s d=$-$0.29), and more negative (curving down, d=$-$0.27) temporal tuning coefficients (Hotelling t-squared test p=0.2). When sounds were added to flashes, however, both transfer functions became less flat (d=0.65 and 0.26 for "Async" compared to “Flash”, for amplitude and temporal tuning respectively), and more like curves for control neurons (Hotelling test p=0.04). Thus, acoustic stimulation tended to tune the network in the same direction as visual stimulation when delivered alone (Figure 2D), but negated the effect of visual stimulation when combined with it. This may imply that multisensory integration in the tectum is dominated by inhibition (see Discussion). Note also that looming stimuli seemed to have weaker effects on neuronal tuning compared to flashes, both in terms of amplitude (d=0.44) and temporal tuning (d=0.22).

Describing neuronal tuning with only a few variables allowed us to compare cell-to-cell variability of tuning in different experimental groups. We found that this variability decreased as neurons were modulated away from the baseline (Bartlett test p=2e-9 for amplitude tuning, p=1e-9 for temporal tuning; Figure 2D). Groups that were significantly different in average values also had different variances (F-test with p$<$0.05), such as Control vs. Flash and Control vs. Looming for both amplitude and temporal tuning. This expands on a finding in our previous study \citep{ciarleglio2015} that prolonged patterned stimulation reduces diversity of tuning profiles in the network, reshaping them according to the spatiotemporal characteristics of the stimulus, with stronger stimuli having stronger effects on tuning diversity.

@@

To see whether prolonged sensory stimulation affected synaptic inputs received by tectal neurons, we recorded evoked excitatory postsynaptic currents in response to optic chiasm stimulation. We found that the amplitude of the early, monosynaptic component of evoked responses (the average current between 5 and 15 ms after the shock; Figure 3B) differed across experimental groups (Figure 3A; F(5,161)=3.2, p=0.009; see Methods for a description of the linear model we used). Both Sync and Async multisensory groups had larger early synaptic currents than the Control group (Tukey p=0.03 and 0.04; Cohen d=0.92 and 0.73 on log-transformed data respectively). The amplitude of late synaptic currents–produced by recurrent network activation (15-145 ms after the stimulus)–did not differ across groups.

@@

% \citep{stemmler1999information}

% , so if all cells from all groups were lumped together in one simple correlation analysis, there was correlation between these two variables (r=$-$0.07, p=0.4, n=151) (

>>>>

It is often assumed, in fields as diverse as connectomics and machine learning, that the main difference between functional and dysfunctional neural networks lies in their connectivity \citep{takemura2014,hildebrand2017,bassett2017network,reimann2017}. However, biological neurons also demonstrate complex and multifaceted intrinsic tuning  \citep{oleary2013,evans2015}, with cells within a single network having different activation thresholds \citep{kole2012}, bursting behaviors \citep{popovic2011}, and inactivation profiles \citep{bianchi2012}. Dysregulation of intrinsic plasticity affects the dynamics of networks  \citep{tien2018}, and can lead to a loss of function \citep{marcelin2009}. And yet, with the possible exception of oscillatory networks \citep{marder2011,picton2018control}, the exact ways in which intrinsic plasticity contributes to network tuning, remain unclear. We know that neurons adjust their spikiness to match the levels of synaptic activation they experience \citep{aizenman2003,titley2017}, but we also know that intrinsic properties affect neuronal tuning to different temporal patterns of activation \citep{azouz2000threshold,branco2010dendritic,fontaine2014threshold,jarvis2018morphology,ohtsuki2018bursting,zbili2019axonnav}. That begs the question: do neurons dynamically adjust this type of tuning in practice?

In this paper, we investigate whether exposure to sensory stimuli with different temporal dynamics would change intrinsic temporal tuning of sensory neurons in the optic tectum of \textit{Xenopus} tadpoles. Tadpole tectum is an ideal model for exploring this question: it is a highly malleable distributed network of about 10\textsuperscript{4} neurons \citep{pratt2013}, involved in stimulus discrimination and sensorimotor transformations \citep{dong2009,khakhalin2014}. In development, tectal neurons acquire diverse phenotypes, that are then shaped by sensory experiences \citep{xu2011,ciarleglio2015}. Circuits in the tectum can learn and reproduce temporal patterns to which they were exposed \citep{pratt2008}: a property that could in principle be achieved through synaptic changes alone \citep{lukovsevivcius2009}, but which may be easier to achieve through intrinsic temporal tuning \citep{narayanan2008,beatty2014}. Finally, tectal neurons exhibit strong Na channel inactivation, which seems to play a role in collision detection \citep{jang2016}, can support temporal tuning \citep{clay2012optimal,fontaine2014threshold,zbili2019axonnav}, and is a known target for plasticity mechanisms \citep{bianchi2012}.

Specifically, we ask three questions about the properties of intrinsic plasticity in tectal networks. First, we test whether the intrinsic temporal tuning to either faster (more synchronous) or slower (asynchronous) inputs would change in response to sensory stimulation. Then, we ask whether intrinsic temporal tuning of individual tectal cells is coordinated with the typical length of synaptic currents they receive. Finally, we try to identify the mechanisms that may underlie temporal tuning variability.

In our previously published large-scale census of tectal cells \citep{ciarleglio2015}, we observed no signs of intrinsic temporal tuning (figures 2, 4), and no interaction between intrinsic and synaptic phenotypes. We argue however, that the current clamp protocols used in earlier studies were not adequate for the task, so in this paper we rely on a dynamic clamp technique. The main benefit of the dynamic clamp, compared to commonly used current step injections, is that it allows a more realistic simulation of neuronal responses to synaptic conductances \citep{prinz2004}. In dynamic clamp, the current injected into the cell is adjusted in real-time, based on a predefined formula that depends on time and cell membrane potential. This means that with dynamic clamp recordings we can excite a neuron in a controlled manner, but still allow both its voltage and transmembrane currents to change, preserving feedback interactions between active currents and transient voltage-gated channels inactivation \citep{ma1996dynamic,zbili2019axonnav}, which is an important mechanism for temporal tuning \citep{branco2010dendritic,platkiewicz2011}. As a second methodological innovation, instead of relying on one type of sensory stimulation \citep{ciarleglio2015}, here we used five different stimulation protocols, and compared to each other their effects on intrinsic tuning.

@@

First we checked whether our stimulation protocols caused any changes in intrinsic excitability of tectal neurons. From previous studies, we knew that in tadpoles exposed to LED flashes, tectal neurons became more excitable \citep{aizenman2003,ciarleglio2015}. However, the stimuli we used in the present study were weaker, and similar to those used in behavioral experiments \citep{khakhalin2014,james2015,truszkowski2017}. We presented a checkerboard pattern that inverted once a second, for four hours; either instantaneously (dubbed “Flash”; Figure 1C left), or with a slow transition over the course of a second (old black squares shrank to white, while new black squares grew from old white squares, dubbed “Looming”; Figure 1C right).

@@

After conditioning, we excised the brain, performed whole cell patch clamp on tectal neurons (Figure 1A), and counted spikes produced in response to dynamically simulated synaptic conductances (Figure 1B). We used conductances of 4 different durations (100, 200, 500, and 1000 ms), and 3 different amplitudes (peak conductances of 0.2, 0.5, and 1.0 nS), matching the range of synaptic currents observed in tectal circuits \textit{in vivo} (\citealt{xu2011,khakhalin2014,ciarleglio2015}; see Methods). Based on prior research \citep{aizenman2003,ciarleglio2015}, we expected neurons exposed to instantaneous checkerboard inversions (flashes) to become more excitable, but surprisingly, they were on average less spiky, and generated 0.4$\pm$0.4 spikes, across all types of dynamic clamp inputs, compared to 0.9$\pm$1.0 spikes in control (Figure 1D, left column; F(1,677)=30.4, p=5e-8, n cells=28, 29). (Here and below, F-values are reported for a multivariate fixed effects analysis of variance with selected interactions, where cell id is used as a repeated measures factor; see Methods for a detailed description). Neurons exposed to looming transitions also spiked less than control neurons (0.6$\pm$0.4 spikes; F(1,629)=9.6; p=0.002; n=28, 25), but more than those exposed to instantaneous "flashes" (F(1,641)=3.7; p=4e-6; n=29, 25). This was likewise unexpected, as compared to flashes, looming stimuli are known to elicit stronger tectal responses \citep{khakhalin2014,khakhalin2019graph}, yet their effect on neuronal excitability was weaker, seemingly violating the principle of homeostatic plasticity \citep{pratt2007,turrigiano2007}.

We then mapped the amplitude tuning of neurons (their amplitude transfer function, or gain), by looking at how an increase in transmembrane conductance translated into increased spike output. Compared to control, neurons from animals exposed to visual stimuli had a flatter amplitude tuning curve (less of a difference between columns of Figure 1D, or between curves in each individual panel of Figure 1E), as they did not increase their spiking as fast in response to larger conductances (F(1,677)=15.9, p=8e-5; and F(1,629)=8.3, p=0.004 for flash- and looming respectively, compared to control). %The flattening of response curves seemed slightly more pronounced for flashes than for looming, but this difference was not significant (F(1,641)=3.5; p=0.06).

These results show that prolonged stimulation had an effect on intrinsic excitability, but its direction was opposite to what was previously described \citep{aizenman2003,ciarleglio2015}, as neurons became less excitable. Moreover, while looming stimuli are known to be more salient, both behaviorally and physiologically \citep{khakhalin2014}, they had a weaker effect on neuronal excitability in comparison to less salient flashes. We conjecture that the difference in the direction of change is due to our stimuli being weaker than those used in earlier studies (see Discussion), and we further explore the difference between flashes and looming stimuli below.

@@

We then examined whether different types of sensory activation would differently reshape temporal intrinsic tuning in tectal neurons. As changes in intrinsic properties seemed homeostatic (increased activation led to reduced spiking), by the same logic, one could expect that exposure to shorter sensory stimuli (flashes; \citealt{khakhalin2014}) would selectively suppress spiking in response to shorter synaptic inputs. Alternatively, shorter stimuli could make neurons better adjusted to working with short bursts of activation \citep{stemmler1999information}, as  described for synaptic processing \citep{aizenman2007} and recurrent activity in the tectum \citep{pratt2007,shen2011}.

We found that compared to control, neurons exposed to flashes and looming stimuli had different response curves when tested across input conductances of different lengths (F(1, 677)=25.1; p=7e-7; and F(1,629)=12.0; p=6e-4 respectively). In control neurons, longer inputs (500 and 1000 ms) typically evoked stronger spiking than shorter inputs (100 and 200 ms), but neurons from stimulated animals had flatter tuning curves, with a plateau, or even a decrease in spike output for longer conductance injections (Figure 1E). In other words, while control neurons preferred longer inputs, stimulated neurons developed a preference for shorter synaptic inputs, and this change was more pronounced in neurons exposed to flashes than in those exposed to looming stimuli (F(1,641)=7.5; p=0.006). This suggests that the change in overall intrinsic excitability, and the change in temporal tuning, follow two different kinds of logic. The overall excitability is homeostatic, as neurons became less excitable in response to stronger stimulation. The temporal retuning however can be better described as “adaptive”, as neurons exposed to shorter stimuli (flashes) became relatively \textit{more} responsive to shorter stimuli, and less responsive to longer stimuli, making them more equipped to process faster, more synchronous patterns of activation \citep{stemmler1999information,fontaine2014threshold}.

@@

While the optic tectum (homologous to superior colliculus in mammals) is often described as a visual area, it is also involved in heavy multisensory computations \citep{stein2014}. In tadpoles, it integrates visual information with inputs from mechanosensory, auditory, and lateral line modalities \citep{deeg2009,pratt2009trigeminal,hiramoto2009,felch2016,truszkowski2017}, but the logic of this integration is unclear. We wondered whether acoustic stimuli would reshape intrinsic properties of tectal neurons, and whether this retuning would be similar to that produced by visual stimuli.

To test this question, we exposed tadpoles to four hours of behaviorally salient, startle-inducing “click” sounds \citep{james2015,truszkowski2017}, provided at the same frequency (every second) as visual stimuli in the first set of experiments. We found (Fig 1D, E) that exposure to these sounds (group “Sound”) did not lead to significant changes in either average number of spikes (0.8$\pm$0.7 spikes; F(1,689)=1.7; p=0.2; n=28, 30), amplitude transfer function (F(1,689)=2.0; p=0.2), or temporal tuning curve (F(1,689)=2.0; p=0.2). This may suggest that acoustic stimuli did not activate tectal circuits strongly enough during conditioning, despite being more behaviorally salient (at the onset of stimulation, acoustic clicks evoked startle responses in about 50-80\% of cases, compared to 5-10\% for checkerboard inversions \citep{james2015,truszkowski2017}). This was not necessarily surprising, as auditory and mechanosensory inputs are known to have different cellular targets in the tectum \citep{pratt2009trigeminal,felch2016,truszkowski2017}, and thus can recruit tectal excitatory and inhibitory networks to a different extent. Alternatively, acoustic stimuli may be inherently weaker in triggering plasticity effects in tectal neurons, as compared to visual inputs, they arrive at different subcellular compartments within the dendritic tree \citep{hiramoto2009,deeg2009}, which can affect their influence on neuronal plasticity \citep{richards2019dendritic}.

We then combined visual and acoustic stimuli in two different ways and looked at the effects of multisensory stimulation on the intrinsic properties of tectal neurons. For some animals, we synchronized the instantaneous checkerboard inversions (flashes) with sound clicks (dubbed “Sync”), while for others, we staggered visual and acoustic stimuli by half a period (500 ms; dubbed “Async”). We found (Fig 1D, E) that multisensory stimulation suppressed excitability of tectal neurons less strongly than visual stimulation alone (0.6$\pm$0.4 spikes, F(1,689)=11.2, p=8e-4; and 0.7$\pm$0.6 spikes, F(1,665)=41.7, p=2e-10, for Sync and Async respectively, compared to Flash, across all testing conditions). Moreover, both temporal and amplitude tuning curves in multisensory groups were less flat than in the Flash group, with a stronger effect in the Async group (for amplitude tuning: F(1,689)=1.3, p=0.3, and F(1,665)=2.4, p=0.02 in Sync and Async groups respectively; for temporal tuning: F(1,689)=9.3, p=0.002, and F(1,665)=22.8, p=2e-6 respectively). This suggests that while on their own sound clicks had little effect on tectal excitability, when added to visual flashes, they negated some of the retuning effects that visual stimulation would have had.

\begin{figure*}[!t]
\includegraphics[width=\linewidth]{fig2.pdf}
\caption{
Quantification of changes in temporal tuning in response to sensory experience. (\textbf{A}). An illustration of how “Temporal tuning” and “Amplitude tuning” values were calculated. For the temporal tuning measure, the value of zero corresponds to linear dependency (blue line), positive values - to an accelerating, supralinear curve (red), and negative values - to plateau- and hill-shaped curves (gray, orange, green). For amplitude tuning, higher values correspond to faster increase in spiking with increased conductance. (\textbf{B}). Amplitude tuning of neurons across different experimental groups (stars show t-test p$<$0.05 compared to control). (\textbf{C}). Temporal tuning and spikiness of neurons in different experimental groups. Neurons from each target group are shown in color, while all neurons from all groups are shown in light gray, as a reference; groups means are shown as black dots; ellipses represent 95\% normal confidence regions. Two outliers (top right corner) are brought within the axes limits. (\textbf{D}). Same data as in (C), shown as averages for each group, with 95\% confidence intervals. Black arrows show the effects of sound clicks, when they were added to control (red), and when they were added to "Flashes" (brown), to form two types of multisensory stimuli.}
\end{figure*}

To better visualize and interpret differences in neuronal tuning after different types of stimulation, we quantified three aspects of intrinsic excitability (average spikiness, amplitude tuning, and temporal tuning) with one value per neuron (see Methods). We used the mean number of spikes across all conditions as the measure of "spikiness"; the linear slope of the number of spikes as a function of conductance amplitude as the measure of "amplitude tuning", and the quadratic term of the curvilinear regression for the number of spikes as a function of input duration as the value to characterize the "temporal tuning" of each neuron (Figure 2A). These "tuning coefficients" capture the character of tuning curves for each neuron (Figure 2A). All three parameters differed across experimental groups: F(5,160)=3.1, p=0.01 for average spikiness (Figure 2C; see Methods for model description); F(5,160)=4.8, p=4e-4 for amplitude tuning (Figure 2B); and F(5,160)=3.6, p=4e-3 for temporal tuning (Figure 2C).

A visual comparison of neuronal tuning in different experimental groups (Fig 2C) shows that acoustic stimulation had opposite effects when provided on its own (without visual stimulation), compared to when it was added to visual flashes. Compared to control neurons, cells exposed to sound clicks seemed to have lower amplitude (Cohen’s d=$-$0.29) and temporal (d=$-$0.27) tuning coefficients (Hotelling t-squared test p=0.2). When sounds were added to flashes, however, both tuning coefficients increased (d=0.65 and 0.26 for amplitude and temporal tuning respectively, "Async" compared to “Flash”), making neurons more similar to control neurons (Hotelling test p=0.04). Thus, acoustic stimulation tended to tune the network in the same direction as visual stimulation when delivered alone (Figure 2D), but it negated the effects of visual stimulation the two were combined. This effect is likely to be either a consequence of strong inhibitory recruitment during multisensory integration in the tectum \citep{liu2016,hamodi2016}, or a sign of highly non-linear interactions between acoustic and visual inputs within the dendritic trees of individual tectal neurons \citep{deeg2011,felch2016,truszkowski2017}.

Somewhat similarly, looming stimuli seemed to have weaker effects on intrinsic properties, compared to flashes, both in terms of amplitude (d=0.44) and temporal tuning (d=0.22), despite them eliciting stronger responses in-vivo \citep{khakhalin2014,khakhalin2019graph}. This suggests that intrinsic retuning depends on stimuli synchronicity, rather than simply the total number of spikes generated by the network, as more synchronous stimuli (Flash, Sync) had stronger effects on tuning than comparable asynchronous stimuli (Async, Looming).

@@

Describing neuronal tuning with only a few variables also allowed us to compare cell-to-cell variability of tuning in different experimental groups. We found that, as it can be guessed from Figure 2D, this variability decreased as neurons were modulated away from the baseline (Bartlett test p=2e-9 for amplitude tuning, p=1e-9 for temporal tuning). Groups that were significantly different in average values also had different variances (F-test with p$<$0.05), such as Control vs. Flash and Control vs. Looming for both amplitude and temporal tuning. This expands on the findings of our previous study \citep{ciarleglio2015}, where we showed that prolonged patterned stimulation reduces diversity of tuning profiles in the network, reshaping them according to the spatiotemporal characteristics of the stimulus. We now show that stimuli that reshape the network stronger, also have a more restrictive effect on tuning diversity.

@@

To see whether prolonged sensory stimulation affected synaptic inputs received by tectal neurons, we recorded evoked excitatory postsynaptic currents in response to optic chiasm stimulation. We found that the amplitude of the early, monosynaptic component of evoked responses (the average current between 5 and 15 ms after the shock; Figure 3B) differed across experimental groups (Figure 3A; F(5,161)=3.2, p=0.009; see Methods for a description of the linear model we used). Both Sync and Async multisensory groups had larger early synaptic currents than the Control group (Tukey p=0.03 and 0.04; Cohen d=0.92 and 0.73 on log-transformed data respectively). The amplitude of late synaptic currents produced by recurrent network activation (15-145 ms after the stimulus) did not differ across groups.

@@
